\section{Vector Spaces}

Let's define a special type of algebraic structure.
\dfn{} {
  A \textbf{vector space} (or \textbf{linear space}) \textit V over a field \textit K consists of a set on which two operations (called \textbf{addition} and \textbf{scalar multiplication}, respectively) are defined so that $\forall x,y \in V$ there is a unique element $x + y \in V$, and $\forall \lambda \in K$ and $\forall x \in V$ there is a unique element $\lambda x \in V$, such that the following conditions hold.
  \begin{itemize}
    \item (VS 1) $\forall x,y \in V,~ x + y = y + x$ (commutativity of addition)
    \item (VS 2) $\forall x,y, z \in V,~ (x + y) + z = x +(y + z)$ (associativity of addition)
    \item (VS 3) $\exists~ \overline 0 \in V$  such that $\forall x \in V,~ x + \overline 0 = x$
    \item (VS 4) $\forall x \in V,~ \exists~ y \in V$ such that $x + y = \overline 0$
    \item (VS 5) $\forall x \in V,~ 1x = x$
    \item (VS 6) $\forall \lambda, \beta \in K,~ x \in V, ~ (\lambda \beta) x = \lambda (\beta x)$
    \item (VS 7) $\forall \lambda \in K,~ x,y \in V,~ \lambda (x + y) = \lambda x + \lambda y$
    \item (VS 8) $\forall \lambda, \beta \in K,~ x \in V,~ (\lambda + \beta)x = \lambda x + \beta x$
  \end{itemize}
}

\nt{
  The element $x + y$ and $\lambda x$ are called \textbf {sum} of \textit x and \textit y, and the \textbf{product} of \textit{$\lambda$} and \textit x, respectively.
  Moreover, the elements of the field \textit K are called \textbf{scalars} and the elements of the vector space \textit V are called \textbf{vectors}.
}

\noindent Note that in describing a vector space, it's necessary to specify not only the vectors but also the operations of addition and scalar multiplication.

We conclude this section with a few of the elementary consequences of the definition of a vector space.

\thm{Cancellation Law for Vector addition}{
  If $x, y, z \in V$ such that $x + z = y + z$, then $x = y$
}
\begin{myproof}
  There exists a vector $v \in V$ such that $z + v = \overline 0$ (VS 4). Thus
  \begin{align*}
    x &= x + \overline 0 = x + (z + v) = (x + z) + v\\
      &= (y + z) + v = y + (z + v) = y + \overline 0 = y
  \end{align*}
  by (VS 2) and (VS 3).
\end{myproof}

\cor{}{
  The vector $\overline 0$ described in (VS 3) is unique.
}

\cor{}{
  The vector \textit y described in (VS 4) is unique.
}

The vector $\overline 0$ in (VS 3) is called the \textbf{zero vector} of \textit V, and the vector \textit y in (VS 4) is called the \textbf{additive inverse} of \textit x and is denoted by $-x$

\thm{}{
  In any vector space \textit V, the following statemets are true:
  \begin{enumerate}[a)]
    \item $\forall x \in V,~0x = \overline 0$
    \item $\forall \lambda \in K,~ x \in V,~ (-\lambda)x = - (\lambda x) = \lambda (-x)$
    \item $\forall \lambda \in K,~ \lambda \overline0 = \overline0$
  \end{enumerate}
}
